### 2021-02-20 paper/mdpi_paper.Rmd
adding rticles formatting

---
title: Full title of the paper (Capitalized)
author:
  - name: Dominik Leutnant
    affil: 1,2,\ddagger,*
    orcid: 0000-0003-3293-2315
  - name: John Doe
    affil: 2, \dagger, \ddagger
affiliation:
  - num: 1
    address: |
      Muenster University of Applied Sciences - 
      Institute for Infrastructure, Water, Resources, Environment
      Correnstr. 25, 48149 Muenster, Germany
    email: leutnant@fh-muenster.de
  - num: 2
    address: |
      Your department
      Street, City, Country
    email: mail@mail.com
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
correspondence: |
  leutnant@fh-muenster.de; Tel.: +XX-000-00-0000.
journal: water
type: article
status: submit
bibliography: mybibfile.bib
appendix: appendix.tex
simplesummary: |
  A Simple summary goes here.
abstract: |
  Canine rabies causes an estimated 60,000 human deaths per year, but these 
  deaths are preventable through post-exposure prophylaxis of people and 
  vaccination of domestic dogs. Dog vaccination campaigns targeting 70% of 
  the population are effective at interrupting transmission. Here, we report on 
  lessons learned during pilot dog vaccination campaigns in the Moramanga 
  District of Madagascar. We compare two different vaccination strategies: 
  a volunteer driven effort to vaccinate dogs in two communes using static 
  point vaccination, and continuous vaccination as part of routine veterinary 
  services. We used dog age data from the campaigns to estimate key demographic 
  parameters and to simulate different vaccination strategies. Overall, 
  we found that dog vaccination was feasible and that most dogs were accessible to vaccination. 
  The static-point campaign achieved higher coverage, but required more 
  resources and had a limited geographic scope compared to the continuous 
  delivery campaign. Our modeling results suggest that targeting puppies 
  through community-based vaccination efforts could improve coverage. We 
  found that mass dog vaccination is feasible and can achieve high coverage 
  in Madagascar, however context-specific strategies and an investment in dog 
  vaccination as a public good will be required to move the country towards 
  elimination. 
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
conflictsofinterest: |
  Declare conflicts of interest or state 'The authors declare no conflict of 
  interest.' Authors must identify and declare any personal circumstances or
  interest that may be perceived as inappropriately influencing the
  representation or interpretation of reported research results. Any role of the
  funding sponsors in the design of the study; in the collection, analyses or 
  interpretation of data in the writing of the manuscript, or in the decision to 
  publish the results must be declared in this section. If there is no role, 
  please state 'The founding sponsors had no role in the design of the study; 
  in the collection, analyses, or interpretation of data; in the writing of the 
  manuscript, an in the decision to publish the results'.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism
output: rticles::mdpi_article
---

# Version

This Rmd-skeleton uses the mdpi Latex template published 2019/02. 
However, the official template gets more frequently updated than the 'rticles'
package. Therefore, please make sure prior to paper submission, that you're 
using the most recent .cls, .tex and .bst files 
(available [here](http://www.mdpi.com/authors/latex)).

# Introduction

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper [@bertrand-krajewski_distribution_1998; @leutnant_stormwater_2016]. And now citing a book reference @gujer_systems_2008. Please use the command [@leutnant_stormwater_2016] for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.

# Materials and Methods

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.

# Results

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.

## Subsection Heading Here

Subsection text here.

### Subsubsection Heading Here

Bulleted lists look like this:

* First bullet
* Second bullet
* Third bullet

Numbered lists can be added as follows:

1. First item
2. Second item
3. Third item

The text continues here.

All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

\begin{figure}[H]
\centering
\includegraphics[width=3 cm]{logo-mdpi}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
entry 1		& data			& data\\
entry 2		& data			& data\\
\bottomrule
\end{tabular}
\end{table}

This is an example of an equation:

\begin{equation}
\mathbb{S}
\end{equation}
<!-- If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph. -->

<!-- Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows: -->
Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

# Discussion

Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

# Conclusion

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

# Patents

This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.